Improving Direct-Control Reinforcement Learning for Network Intrusion Prevention (WIP)

### 2018-08-31 scone-talk/scone-talk.tex

@@

@@

%bib
\usepackage[maxnames=3,maxbibnames=99,mincrossrefs=5,sortcites
,backend=bibtex
,style=authortitle
]{biblatex}
\addbibresource{papers-off.bib}
\addbibresource{confs-off.bib}
\addbibresource{books-off.bib}
\addbibresource{rfc.bib}
\addbibresource{misc.bib}

@@

\title{Reinforcement Learning in Network Defence/Control (WIP)}
\author{Kyle A. Simpson\\
	\small{\faGithub{} \href{https://github.com/felixmcfelix}{FelixMcFelix} \hspace{0.5em} \faGlobe{} \url{https://mcfelix.me}}}
\institute{University of Glasgow}
\date{\nth{10} April, 2018}

@@

\section{Reinforcement learning}

\begin{frame}{Introducing: Reinforcement Learning (RL)}
	\begin{itemize}
		\item Goal: train an agent to make optimal decisions based on observed state.
		\begin{itemize}
			\item Formally, maximise the \alert{expected discounted reward}\footcite{RL2E}.
		\end{itemize}

		\item Underlying theory: systems as (discrete-time) \alert{Markov Decision Processes}---states, actions, rewards and transition probabilities.
		\begin{itemize}
			\item I.e., choosing action $a_t$ from a policy in state $s_t$, $a_t \sim \pi(s_t)$, induces the next state $s_{t+1}$ and an associated reward $r_{t+1}$ stochastically.
			\item Generalises to \alert{value} $Q(s,a)$---how much reward can we \emph{eventually} expect from choosing each action currently available?
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Why have RL in networks?}
	%Link to recent work in data-driven networking/optimisation?
	%
	%Link to that one GMM paper w/ RL communication.

	\begin{itemize}
		\item \alert{Data-driven networking.} Effectively applied to intra-domain routing \footcite{DBLP:conf/hotnets/ValadarskySST17}, task allocation \footcite{DBLP:conf/hotnets/MaoAMK16} and more, each with general and domain-specific insights.

@@

\section{What I'm working on}

@@

		\item Strong assumptions on benign/malicious packet estimation quality. 
		\item Pushback is known to be flawed in practice.
		\item Rigid ISP-like network assumptions.
	\end{itemize}
\end{frame}

@@

			\item More granular $\implies$ should focus on flow stats, \alert{not aggregates} like now! %Alongside this higher-granularity view, reformulation to include flow characteristics in the decision-making process.
		\end{itemize}
	\end{column}
	\begin{column}{0.5\linewidth}
		\begin{figure}
			\includegraphics[width=\linewidth]{../plots/online-pres.pdf}
			\caption{Service quality decreases as actions become less granular.}
		\end{figure}
	\end{column}
	\end{columns}
\end{frame}

\section{The adversarial setting}

\begin{frame}{Attacks}
	\begin{description}
		\item[Poisoning attacks] Affect the training of a continuously learning model (RL agents, collaborative learning).

		\item[Evasion attacks] Make data/state appear to be different. These are what are often known as \alert{adversarial examples}!
	\end{description}
\end{frame}

\begin{frame}{The capabilities of attackers}
	\begin{itemize}
		\item \emph{Polymorphic blending attacks} (i.e., evasion) on past models are known \alert{\NP-Complete} by reduction to (s)FSA\footcite{DBLP:conf/ccs/FoglaL06}.

		\item We have empirical bounds on the effectiveness of poisoning attacks wrt.\ what percentage of inbound traffic an attacker controls\footcite{DBLP:journals/jmlr/KloftL10}.

		\item Careless model exposure $\implies$ \alert{black-box model extraction}\footcite{DBLP:conf/uss/TramerZJRR16} $\implies$ evasion\footcite{DBLP:conf/sp/Carlini017}.

		%\item Promising work on defences\footcite{DBLP:conf/acsac/CaoG17}.
	\end{itemize}
\end{frame}

\begin{frame}{Unknowns}
	\begin{itemize}
		\item We don't have results for poisoning on e.g., Deep RL in the network.
		\begin{itemize}
			\item Can an adversary precisely affect state at all measurement points?
			% Networks are unstable ``bursty'', and attacker has limited control over inbound traffic.
		\end{itemize}

		\item We also lack results on whether evasion attacks can transfer between policies which evolve over time.
		\begin{itemize}
			\item E.g., $\pi_t$ to $\pi_{t+\delta}$ for reinforcement learners focused on non-stationarity.
		\end{itemize}

	\end{itemize}
\end{frame}

\begin{frame}[standout]{Conclusion}
	We've looked at...
	\begin{itemize}
		\item A quick introduction to RL, and its \alert{importance to future networks}.
		\item A `direct control' approach to intrusion prevention, and \alert{intended improvements}.
		\item Adversarial work, security, and open questions.
	\end{itemize}

>>>>

% !TeX program = lualatex

@@

%\usepackage[T1]{fontenc}

@@

%bib
\usepackage[maxnames=3,maxbibnames=99,mincrossrefs=5,sortcites
,backend=bibtex
,style=authortitle
]{biblatex}
\addbibresource{../paper/papers-off.bib}
\addbibresource{../paper/confs-off.bib}
\addbibresource{../paper/books-off.bib}
\addbibresource{../paper/rfc.bib}
\addbibresource{../paper/misc.bib}

@@

\title{Improving Direct-Control Reinforcement Learning for Network Intrusion Prevention (WIP)}
\author{Kyle A. Simpson\\
	\small{\faGithub{} \href{https://github.com/felixmcfelix}{FelixMcFelix} \hspace{0.5em} \faGlobe{} \url{https://mcfelix.me}}}
\institute{University of Glasgow}
\date{\nth{3} September, 2018}

@@

\begin{frame}{Introduction}
	\begin{itemize}
		\item Network IDS/IPS backed by machine learning haven't taken off as hoped---particularly anomaly-based work.
		\item Detection problem tricky in this domain:
		\begin{itemize}
			\item Evolving: usage shifts, new protocols, new applications.
			\item Burstiness, seasonal variation.
			\item Need for correctness, almost no false-positive tolerance.
			\item Labelling issues.
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Introduction: Part II}
\begin{itemize}
	\item Classes of problem like flooding-based DDoS attacks manifest as a service degradation.
	\begin{itemize}
		\item Can these be controlled via feedback loop?
		\item \alert{``Overcome'' the difficulties of the detection problem} by monitoring and adapting to \emph{performance characteristics and consequences} in real-time
	\end{itemize}
	\item Goal: augment signature-based approaches as a last-ditch effort.
\end{itemize}
\end{frame}

%\section{Reinforcement learning}

\begin{frame}{RL: The Main Idea\texttrademark}
	\begin{itemize}
		\item Goal: train an agent to make optimal decisions based on observed state.
		\begin{itemize}
			\item Formally, learn a \alert{policy} to maximise the \alert{expected discounted reward}\footcite{RL2E}.
		\end{itemize}

		\item Underlying theory: systems as (discrete-time) \alert{Markov Decision Processes}---states, actions, rewards and transition probabilities.
		\begin{itemize}
			\item I.e., choosing action $a_t$ from a policy in state $s_t$, $a_t \sim \pi(s_t)$, induces the next state $s_{t+1}$ and an associated reward $r_{t+1}$.
			\item Generalises to \alert{value} $Q(s,a)$---how much reward can we \emph{eventually} expect from choosing each action currently available?
		\end{itemize}

%		\item The world/environment is believed to be modelled by stochastic state-transition probabilities, reward function distributions...
%		\begin{itemize}
%			\item But we don't need to model that!
%		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{RL: The Main Benefits\texttrademark}
\begin{itemize}
	\item We can learn the optimal policy \alert{without modelling the world ourselves}.
	\item Formulation allows learning adaptively and online, so long as a reward signal is available.
	\item Variation in available algorithms, update mechanisms, function approximations, dependence on value functions, action selection, exploration...
	\begin{itemize}
		\item Orthogonal concerns, allowing tunable algorithm design.
	\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Where has RL succeeded in networks?}
	%Link to recent work in data-driven networking/optimisation?
	%
	%Link to that one GMM paper w/ RL communication.

	\begin{itemize}
		\item \alert{Data-driven networking.} Effectively applied to intra-domain routing \footcite{DBLP:conf/hotnets/ValadarskySST17}, task allocation \footcite{DBLP:conf/hotnets/MaoAMK16}, traffic optimisation \footcite{DBLP:conf/sigcomm/ChenL0L18} and more, each with general and domain-specific insights.

@@

\begin{frame}{What might we gain from RL in DDoS prevention?}
content...
\end{frame}

%\section{Existing work}

@@

		\item Strong assumptions on benign/malicious packet estimation quality. 
		\item Pushback is known to be flawed in practice.
		\item Rigid ISP-like network assumptions.

		\item FIXME WITH THE NEW GRIPES
	\end{itemize}
\end{frame}

@@

			\item More granular $\implies$ should focus on flow stats, \alert{not aggregates} like now! %Alongside this higher-granularity view, reformulation to include flow characteristics in the decision-making process.
		\end{itemize}
	\end{column}
	\begin{column}{0.5\linewidth}
		\begin{figure}
%			\includegraphics[width=\linewidth]{../plots/online-pres.pdf}
			\caption{Service quality decreases as actions become less granular. ??FIXME: MISSING PLOT}
		\end{figure}
	\end{column}
	\end{columns}
\end{frame}

\begin{frame}{The Immediate Future}
content...
\end{frame}

\begin{frame}{The Far Future}
\begin{itemize}
	\item Other problems.
	\begin{itemize}
		\item New action spaces, careful consideration.
	\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[standout]{Conclusion}
	We've looked at...
	\begin{itemize}
		\item A quick introduction to RL, and its \alert{importance to future networks}.
		\item A `direct control' approach to intrusion prevention, and \alert{intended improvements}.
		\item ?? FIXME
	\end{itemize}
